---
title: "Data_Processing_and_Ranking_comparison"
author: "Jiayi Cui, jc4884"
date: "October 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages and data
```{r,warning=FALSE,message=FALSE}
packages.used=c('dplyr', 'geosphere', 'zipcode', 'tigris', 'sp','maptools','broom','httr','rgdal', 'ggmap')
# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}
library(dplyr)
library(geosphere)
library(zipcode)
library(tigris)
library(sp)
library(maptools)
library(broom)
library(httr)
library(rgdal)
source("../lib/na_process.R")
```
```{r}
rest_raw <- read.csv('../data/DOHMH_New_York_City_Restaurant_Inspection_Results.csv')
alcohol <- read.csv('../data/Liquor_Authority_Quarterly_List_of_Active_Licenses.csv')
bus_raw <- read.csv("../data/bus_stop.csv")
subway_raw <-  read.csv("../data/subwayinfo.csv")
price_raw <- read.csv('../data/Zip_Zri_AllHomesPlusMultifamily.csv')
medical_raw <- read.csv('../data/NYC_Health_and_Hospitals_Corporation_Facilities.csv')
market_raw <- read.csv('../data/Retail_Food_Stores.csv')
theatre_raw <- read.csv('../data/DOITT_THEATER_01_13SEPT2010.csv')
crime_raw <- read.csv('../data/NYPD_Complaint_Data_Current_YTD.csv')
gallery_raw <- read.csv('../data/ART_GALLERY.csv')
boro_zip = read.csv("../data/manhattan_zipcode.csv",as.is = T)
save(boro_zip ,file="../output/boro_zip.Rdata")
park_raw <- read.csv("../data/OpenData_ParkZones.csv",as.is = T)
carcrash_raw <- read.csv("../data/NYPD_Motor_Vehicle_Collisions.csv", as.is = T)
comp311_raw <- read.csv("../data/complaints311_processed.csv", as.is = T)
all_other_raw <- read.csv("../data/all_other.csv", as.is = T)
housing_raw0 <- read.csv("../output/new_housing.csv", as.is = T)
housing_raw0 <- cbind(housing_raw0, X = 1:nrow(housing_raw0))
housing_raw <- housing_raw0[,c(9,8,1,2,3,4,5,6,7)] %>% arrange(zipcode)
```

# Match crime location with zipcode
```{r}
zip = paste(boro_zip[,2], collapse = " ")
zip = gsub(",", "", zip)
zip.num = as.numeric(strsplit(zip, " ")[[1]])

crime_manh <- crime_raw %>%
  filter(BORO_NM == 'MANHATTAN')

zipcode_man <- c(seq(10001, 10007), seq(10009, 10014), seq(10016, 10019), seq(10021, 10040), 
             "10065", "10069", "10075", "10128", "10280", "10282", "11106", "11109", "11231")
zip_df <- data.frame(zip_man = zipcode_man)


data("zipcode")
man_zip <- zipcode %>%
  filter(zip %in% zipcode_man)

# crime location and zipcode distance matrix
dist_mat <- distm(crime_manh[,c('Longitude','Latitude')], man_zip[,c('longitude','latitude')],
             fun=distVincentyEllipsoid)
# match by shortest distance
crime_manh$zipcode <- man_zip$zip[max.col(-dist_mat)]

write.csv(crime_manh, file = "../output/Crime_manhattan.csv")
```


# Data processing

## Remove wrong data

Some houses' longitude and latitude are clearly inconsistent with their zipcode or address. Remove these rows.

```{r}
housing_raw = housing_raw %>% filter(lat<=40.9 & lat>=40.7 & lng>=-74.1 & lng<=-73.8)
```

## Prepare dataframe

```{r}
rest_manh_v <- rest_raw %>%
  distinct(BUILDING, STREET, ZIPCODE,.keep_all = TRUE) %>%
  filter(BORO == "MANHATTAN") %>%
  filter(ZIPCODE %in% zipcode_man) %>%
  select(DBA, BUILDING, STREET, ZIPCODE, CUISINE.DESCRIPTION)

alcohol <- alcohol %>%
  distinct(License.Serial.Number, .keep_all = TRUE) %>%
  filter(County.Name..Licensee. == "NEW YORK")

bar <- alcohol %>%
  filter(Zip %in% zipcode_man) %>%
  filter(License.Type.Name == "ON-PREMISES LIQUOR")

club <- alcohol %>%
  filter(Zip %in% zipcode_man) %>%
  filter(License.Type.Name == "CLUB LIQUOR" | License.Type.Name == "CLUB WINE")

bus <- bus_raw %>%
  filter(zipcode %in% zipcode_man)
  
subway <- subway_raw %>%
  filter(zipcode %in% zipcode_man)

hospital <- medical_raw %>%
  filter(Postcode %in% zipcode_man)

market <- market_raw %>%
  filter(Zip.Code %in% zipcode_man)

theatre <- theatre_raw %>%
  filter(ZIP %in% zipcode_man)

gallery <- gallery_raw %>%
  filter(ZIP %in% zipcode_man)

price <- price_raw %>%
  filter(RegionName %in% zipcode_man)

park <- park_raw %>%
  filter(ZIPCODE %in% zipcode_man)

carcrash <- carcrash_raw %>%
  filter(ZIP.CODE %in% zipcode_man)

complaint <- comp311_raw %>%
  filter(Incident.Zip %in% zipcode_man) %>% filter(as.numeric(substr(Created.Date,7,10)) >= 2018)

housing <- housing_raw %>%
  filter(zipcode %in% zipcode_man)
```

# Count by zipcode, fill in NAs, rank and calculate probability

If the count of a zipcode area is missing, fill it with the nearest 6 zipcode area to it.

The probability is the CDF of a uniform distribution from 0 to the maximum number in count.

```{r}
# restaurant violation
count.restaurant_violation <- rest_manh_v %>%
  group_by(ZIPCODE) %>%
  summarise(COUNT = n()) %>% 
  merge(zip_df, by.x = "ZIPCODE", by.y = "zip_man", all.y = T)
for (i in which(is.na(count.restaurant_violation)[,2])){
  count.restaurant_violation[i,2] = na_process(count.restaurant_violation, i)
}
count.restaurant_violation <- count.restaurant_violation %>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))
  
# american style restaurant
count.rest_american <- all_other_raw[,c("zipcode","count.American")] %>%
  merge(zip_df, by.x = "zipcode", by.y = "zip_man", all.y = T)
for (i in which(is.na(count.rest_american)[,2])){
  count.rest_american[i,2] = na_process(count.rest_american, i)
}
count.rest_american <- count.rest_american %>% 
  mutate(RANKING = rank(desc(count.American))) %>%
  mutate(PROB = punif(count.American, max = max(count.American), min = 0))

# chinese style restaurant
count.rest_chinese <- all_other_raw[,c("zipcode","count.Chinese")] %>%
  merge(zip_df, by.x = "zipcode", by.y = "zip_man", all.y = T)
for (i in which(is.na(count.rest_chinese)[,2])){
  count.rest_chinese[i,2] = na_process(count.rest_chinese, i)
}
count.rest_chinese <- count.rest_chinese %>% 
  mutate(RANKING = rank(desc(count.Chinese))) %>%
  mutate(PROB = punif(count.Chinese, max = max(count.Chinese), min = 0))

# italian style restaurant
count.rest_italian <- all_other_raw[,c("zipcode","count.Italian")] %>%
  merge(zip_df, by.x = "zipcode", by.y = "zip_man", all.y = T)
for (i in which(is.na(count.rest_italian)[,2])){
  count.rest_italian[i,2] = na_process(count.rest_italian, i)
}
count.rest_italian <- count.rest_italian %>% 
  mutate(RANKING = rank(desc(count.Italian))) %>%
  mutate(PROB = punif(count.Italian, max = max(count.Italian), min = 0))

# japanese style restaurant
count.rest_japanese <- all_other_raw[,c("zipcode","count.Japanese")] %>%
  merge(zip_df, by.x = "zipcode", by.y = "zip_man", all.y = T)
for (i in which(is.na(count.rest_japanese)[,2])){
  count.rest_japanese[i,2] = na_process(count.rest_japanese, i)
}
count.rest_japanese <- count.rest_japanese %>% 
  mutate(RANKING = rank(desc(count.Japanese))) %>%
  mutate(PROB = punif(count.Japanese, max = max(count.Japanese), min = 0))

# pizza restaurant
count.rest_pizza <- all_other_raw[,c("zipcode","count.Pizza")] %>%
  merge(zip_df, by.x = "zipcode", by.y = "zip_man", all.y = T)
for (i in which(is.na(count.rest_pizza)[,2])){
  count.rest_pizza[i,2] = na_process(count.rest_pizza, i)
}
count.rest_pizza <- count.rest_pizza %>% 
  mutate(RANKING = rank(desc(count.Pizza))) %>%
  mutate(PROB = punif(count.Pizza, max = max(count.Pizza), min = 0))

# bar
count.bar <- bar %>%
  group_by(Zip) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "Zip", by.y = "zip_man", all.y = T)
for (i in which(is.na(count.bar)[,2])){
  count.bar[i,2] = na_process(count.bar, i)
}
count.bar <- count.bar %>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# club
count.club <- club %>%
  group_by(Zip) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "Zip", by.y = "zip_man", all.y = T)
count.club <- new_df(count.club)
count.club <- count.club %>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# subway
count.subway <- subway %>%
  group_by(zipcode) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "zipcode", by.y = "zip_man", all.y = T)
count.subway <- new_df(count.subway) %>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# bus
count.bus <- bus %>%
  group_by(zipcode) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "zipcode", by.y = "zip_man", all.y = T)
count.bus <- new_df(count.bus) %>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# hospital
count.hospital <- hospital %>%
  group_by(Postcode) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "Postcode", by.y = "zip_man", all.y = T) %>%
  new_df()%>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# market
count.market <- market %>%
  group_by(Zip.Code) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "Zip.Code", by.y = "zip_man", all.y = T) %>%
  new_df()%>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# theatre
count.theatre <- theatre %>%
  group_by(ZIP) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "ZIP", by.y = "zip_man", all.y = T) %>%
  new_df()%>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# gallery
count.gallery <- gallery %>%
  group_by(ZIP) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "ZIP", by.y = "zip_man", all.y = T) %>%
  new_df()%>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# crime
count.crime <- crime_manh %>%
  group_by(zipcode) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "zipcode", by.y = "zip_man", all.y = T) %>%
  new_df()%>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# park
count.park <- park %>%
  group_by(ZIPCODE) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "ZIPCODE", by.y = "zip_man", all.y = T) %>%
  new_df()%>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# car crash
count.carcrash <- carcrash %>%
  group_by(ZIP.CODE) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "ZIP.CODE", by.y = "zip_man", all.y = T) %>%
  new_df()%>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))

# complaint
count.complaint <- complaint %>%
  group_by(Incident.Zip) %>%
  summarise(COUNT = n()) %>%
  merge(zip_df, by.x = "Incident.Zip", by.y = "zip_man", all.y = T) %>%
  new_df()%>%
  mutate(RANKING = rank(desc(COUNT))) %>%
  mutate(PROB = punif(COUNT, max = max(COUNT), min = 0))
```

# Output individual features
```{r}
# restaurants "restaurant_overall"
save(count.restaurant_violation,file="../output/count_restaurant_violation.Rdata")
save(count.rest_american,file="../output/rest_american.Rdata")
save(count.rest_chinese,file="../output/rest_chinese.Rdata")
save(count.rest_italian,file="../output/rest_italian.Rdata")
save(count.rest_japanese,file="../output/rest_japanese.Rdata")
save(count.rest_pizza,file="../output/rest_pizza.Rdata")
# party "party"
save(count.bar,file="../output/count_bar.Rdata")
save(count.club,file="../output/count_club.Rdata")
# transportation "travel"
save(count.subway,file="../output/count_subway.Rdata")
save(count.bus,file="../output/count_bus.Rdata")
save(count.carcrash,file="../output/count_carcrash.Rdata")
# art "entertainment"
save(count.theatre,file="../output/count_theatre.Rdata")
save(count.gallery,file="../output/count_gallery.Rdata")
# single features
save(count.hospital,file="../output/count_hospital.Rdata")
save(count.market,file="../output/count_market.Rdata")
save(count.crime,file="../output/count_crime.Rdata")
save(count.park,file="../output/count_park.Rdata")
save(count.complaint,file="../output/count_complaint.Rdata")
```   

# Join features
```{r}
travel <- cbind(count.bus, count.subway[,-1])
colnames(travel) = c("zip",paste(rep(c("bus", "subway"), each = 3),rep(c("count","rank","prob"),2),sep = "."))

party <- cbind(count.bar, count.club[,-1])
colnames(party) = c("zip",paste(rep(c("bar", "club"), each = 3),rep(c("count","rank","prob"),2),sep = "."))

entertainment <- cbind(count.theatre, count.gallery[,-1])
colnames(entertainment) = c("zip",paste(rep(c("theatre", "gallery"), each = 3),rep(c("count","rank","prob"),2),sep = "."))

restaurant_overall <- cbind(count.rest_american, count.rest_chinese[,-1], count.rest_italian[,-1], count.rest_japanese[,-1], count.rest_pizza[,-1], count.restaurant_violation[,-1])
colnames(restaurant_overall) = c("zip",paste(rep(c("american", "chinese", "italian", "japanese", "pizza", "violation"), each = 3),rep(c("count","rank","prob"),5),sep = "."))

feature_names = c("restaurant_violation", "rest_american", "rest_chinese", "rest_italian", "rest_japanese", "rest_pizza", "bar","club","subway","bus","hospital","market","theatre","gallery","crime","park","carcrash","complaint")
feature_all = cbind(as.numeric(zipcode_man), count.restaurant_violation[,-1], count.rest_american[,-1], count.rest_chinese[,-1], count.rest_italian[,-1], count.rest_japanese[,-1], count.rest_pizza[,-1], count.bar[,-1], count.club[,-1], count.subway[,-1], count.bus[,-1], count.hospital[,-1], count.market[,-1], count.theatre[,-1], count.gallery[,-1], count.crime[,-1], count.park[,-1], count.carcrash[,-1], count.complaint[,-1])
colnames(feature_all) <- c("zipcode", paste(rep(feature_names, each = 3),rep(c("count","rank","prob"),18)))

# combine housing info and features into 1 dataframe
housing_all <- merge(housing, feature_all, by.x = "zipcode", by.y = "zipcode")

# the zipcode that do not have any houses list:
zip_none = setdiff(zipcode_man, unique(housing$zipcode))

# save into RData
save(travel,file="../output/travel.Rdata")
save(party,file="../output/party.RData")
save(entertainment,file="../output/entertainment.RData")
save(restaurant_overall,file="../output/restaurant_overall.Rdata")
save(feature_all,file="../output/feature_all.RData")
save(housing_all,file="../output/housing_all.RData")
save(zip_none,file="../output/zip_none.RData")

# names(count.theatre)[names(count.theatre) == 'ZIP'] <- 'Zip'
# names(count.gallery)[names(count.gallery) == 'ZIP'] <- 'Zip'
# count.gallery$Zip <- as.factor(count.gallery$Zip)
# count.theatre$Zip <- as.factor(count.theatre$Zip)
# 
# entertainment <- count.bar %>%
#   full_join(count.club, by = 'Zip') %>%
#   full_join(count.gallery, by = 'Zip') %>%
#   full_join(count.theatre, by = 'Zip')
# count.restaurant_violation$ZIPCODE = as.numeric(count.restaurant_violation$ZIPCODE)
# restaurant_overall <- count.rest_american %>%
#   full_join(count.rest_chinese, by = 'zipcode') %>%
#   full_join(count.rest_italian, by = 'zipcode') %>%
#   full_join(count.rest_japanese, by = 'zipcode') %>%
#   full_join(count.rest_pizza, by = 'zipcode') %>%
#   full_join(count.restaurant_violation, by = c("zipcode"="ZIPCODE"))

```


# Output processed data
```{r}
write.csv(travel, '../output/travel_count.csv')
write.csv(entertainment, '../output/entertainment_count.csv')
write.csv(count.crime, '../output/crime_count.csv')
write.csv(count.market, '../output/market_count.csv')
write.csv(count.hospital, '../output/hospital_count.csv')
write.csv(count.restaurant_violation, '../output/restaurant_violation_count.csv')
write.csv(price, '../output/median_price_manhattan.csv')
```

